---
title: "Homework 1"
author: "Veronica Hurtado"
date: '2019-05-11'
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# PART III

Defining the likelihood, prior and posterior.

```{r}
likelihood<-rpois(10, 1)
alpha_0=1
beta_0=1

prior<-rgamma(10, alpha_0, beta_0)
hist(prior, main="Histogram of the Prior Distribution")

posterior<-likelihood*prior
posterior1<-as.data.frame(posterior)
```

Ploting the posterior distribution as a histogram 

```{r}
hist(posterior, main="Histogram of the Posterior Distribution", xlab="Theta", ylab="Frequency", border="blue", col="darkblue")
d<-density(posterior)
```

Attempt with ggplot2

```{r}
library(ggplot2)
p<-ggplot(posterior1, aes(x=posterior)) +  geom_histogram(binwidth=0.5, color="black", fill="lightblue") 
p + ggtitle("Histogram of the posterior distribution") + ylab("Density") + xlab("Posterior Distribution") + theme(legend.position = "top" )
```

Density plot with credibility intervals and mean.

```{r}
alpha_1=sum(posterior, na.rm = FALSE)+alpha_0
alpha_1

beta_1=beta_0+10
beta_1

cred<-qgamma(c(0.025,0.975), shape=alpha_1, rate=beta_1)
```

Plotting density funciton with relevant markers

```{r}
mean(posterior)

plot(d, main="Posterior Distribution", xlab="Theta", ylab="Probability")
abline(v=mean(posterior), col="red", lty=1) 
text(mean(posterior)+1,0.1,labels="Expected Mean", cex=1)

cred[1]

abline(v=cred[1], col="blue", lty=3)
text(cred[1]-1,0.25,labels="Lower limit", cex=1)

cred[2]

abline(v=cred[2], col="blue", lty=3)
text(cred[2]+1,0.25,labels="Upper limit", cex=1)
```

# PART IV

Repeat with 1000 random numbers from the poisson distribution. 

Defining the likelihood, prior and posterior.

```{r}
likelihood2<-rpois(1000, 1)

prior2<-rgamma(1000, alpha_0, beta_0)
hist(prior2, main="Histogram of the Prior Distribution")

posterior2<-likelihood2*prior2
posterior2_<-as.data.frame(posterior2)
```

Ploting the posterior distribution as a histogram 

```{r}
hist(posterior2, main="Histogram of the Posterior Distribution", xlab="Theta", ylab="Frequency", border="blue", 
     col="darkblue")
d2<-density(posterior2)
```

Density plot with credibility intervals and mean.

```{r}
alpha_1=sum(posterior2)+alpha_0
alpha_1

beta_1=beta_0+1000
beta_1

cred2<-qgamma(c(0.025,0.975), shape=alpha_1, rate=beta_1)
cred2

cred2[1]
cred2[2]

```

Plotting density funciton with relevant markers

```{r}
plot(d2, main="Posterior Distribution", xlab="Theta", ylab="Probability")
abline(v=mean(posterior2), col="red", lty=1)
text(mean(posterior2)+1,0.1,labels="Expected Mean", cex=1)
abline(v=cred2[1], col="blue", lty=3)
text(cred2[1]-0.75,0.25,labels="Lower limit", cex=1)
abline(v=cred2[2], col="blue", lty=3)
text(cred2[2]+3,0.25,labels="Upper limit", cex=1)


```
