---
title: "POLI574 Homework 1"
author: "Nathan Peng"
date: "9 May 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(varbvs)
```

#PART III

```{r}
#FOr 2.2

#Setting Parameters and Generating Data
a1=1
b1=1
set.seed(1)
lik1<-rpois(10, 1)
set.seed(2)
pri1<-rgamma(10, shape=a1, rate=b1)
post1<-lik1*pri1

#Note: I use set.seed to make labeling for the density plot easier.

#Plotting Histogram
d1<-density(post1)
hist(post1, main="Histogram of Posterior Distribution", xlab="Theta", ylab="Density", freq=F)

```

```{r}
#I used the posterior definitions for aplha and beta to generate the credibility intervals (apha1 = Sigma(yi) + alpha0, and beta1=n + beta0), using the gamma quantile (qgamma) command

a0=11+1
b0=10+1
quant1<-cred(post1, mean(post1), w=NULL, cred.int=0.95)

#Plotting density funciton with relevant markers

plot(d1, main="Posterior Distribution", xlab="Theta", ylab="Density")
abline(v=mean(post1), col="red", lty=1)
text(mean(post1)+0.5,0.6,labels="Expected Mean", cex=1)

abline(v=quant1$a, col="blue", lty=3)
text(quant1$a-0.5,0.25,labels="Lower limit", cex=1)

abline(v=quant1$b, col="blue", lty=3)
text(quant1$b+0.5,0.25,labels="Upper limit", cex=1)

```


#PART IV

```{r}
#Setting Parameters and Generating Data
set.seed(1)
lik2<-rpois(1000, 1)
set.seed(2)
pri2<-rgamma(1000, shape=a1, rate=b1)
post2<-lik2*pri2

#Note: I use set.seed to make labeling for the density plot easier.


#Plotting Histogram
hist(post2, main="Histogram of Posterior Distribution, n=1000", xlab="Theta", ylab="Density", freq=F)
d2<-density(post2)
```

```{r}
#Again, I used the posterior definitions for aplha and beta to generate the credibility intervals (apha1 = Sigma(yi) + alpha0, and beta1=n + beta0), using the gamma quantile (qgamma) command
a2=sum(post2)+1
b2=1000+1
quant2<-cred(post2, mean(post2), w=NULL, cred.int=0.95)

#Plotting density funciton with relevant markers
plot(d2, main="Posterior Distribution", xlab="Theta", ylab="Density", xaxp  = c(-1, 10, 11))
abline(v=mean(post2), col="red", lty=1)
text(mean(post2)+3,0.4,labels="Expected Mean", cex=0.75)

abline(v=quant2$a, col="blue", lty=3)
text(quant2$a+1.5,0.25,labels="Lower limit", cex=0.75)

abline(v=quant2$b, col="blue", lty=3)
text(quant2$b+1.5,0.25,labels="Upper limit", cex=0.75)
```

# Comparing the two graphs in Parts III and IV, we see that as n gets large, the mean of the posterior results approximate the data over the prior - the mean of the (i) prior and the (ii) data drawn from the poisson distribution in III is 0.3844242 and 1.1, and for PART IV the prior and data means were 0.992095 and 1.015 in IV. I also note that that the width of the credibility interval does not necessarily decrease with n as the density distribution of the posterior got wider in IV with more values being drawn. 
