---
title: "Homework 2"
author: "Veronica Hurtado"
date: '2019-05-19'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PART II

3.1. I plot the two beta distributions using the parameters gather for n=100 and n=1000.  

```{r }

d1 <- rbeta(100, 21, 91)
d2 <- rbeta(1000, 192, 820)

plot(density(x=d2), col = "red",  main="Density plot for the posterior distribution.", sub="Red=1000, Blue=100.")
lines(density(x=d1), col = "blue")
```

3.8. Bayes Facto for Sam and Levi

```{r}

sam <- c(0.9, 0.8, 0.7, 0.6, 0.5)
levi <- c(0.1, 0.2, 0.3, 0.4, 0.5)

e1 = 10
e2 = 20

m1 <- 100
m2 <- 200

M1a <- gamma(e1+1) / (2*m1 +1)^(e1+1)
M2a <- (gamma(e1*sam +1)/(m1+1)^(e1*sam+1)) * (gamma(e1*levi +1)/(m1+1)^(e1*levi +1)) 

BF_1 <- M1a/M2a

l1 <- log10(BF_1)

M1b <- gamma(e2+1) / (2*m2 +1)^(e2+1)
M2b <- (gamma(e2*sam +1)/(m2+1)^(e2*sam+1)) * (gamma(e2*levi +1)/(m2+1)^(e2*levi +1)) 

BF_2 <- M1b/M2b

l2 <- log10(BF_2)

l1
l2

```

```{r}


```

## PART III

6.1,

```{r}
dat <- c(1/3, 2/3, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 1/4, 0, 0, 1/4, 1/2, 0, 0, 0, 0, 1/8, 7/8, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 1, 0)

P <- matrix(data = dat, nrow = 6, ncol = 6, byrow = TRUE)

library(matrixcalc)
matrix.power(P,100)

matrix.power(P,101)
```

6.3. 

a)

```{r}

data <- c(0.75, 0.25, 0.125, 0.875)
P_1 <- matrix(data = data, nrow = 2, ncol = 2, byrow = TRUE)
matrix.power(P_1,2)

```

b)

I take the vector of states (1,2) and sample one of those. Then, I create a loop for "burnin". Inside the loop, if state ==1, I ask the program to do a draw from the binomial distribution. Here, the third paramter is the specification that the second colum of the first row is the prob of changing from state 1 to state 2. Draw =1 is a success from the binomial distirbution. Success means that you change the state. 


```{r}

# Create MCMC function:
mcmc <- function(p, states, burnin, runs) {
    state <- sample(range, 1)
    for (i in 1:burnin) {
        if (state == 1) {
            draw <- rbinom(1, 1, p[1, 2])
            if (draw == 1) {
                state <- 2 
            } 
        } 
        if (state == 2) {
            draw <- rbinom(1, 1, p[2, 1])
            if (draw == 1) {
                state <- 1
            }
        }
    }
    # storing results
    states <- rep(NA, runs)
    for (i in 1:runs) {
        if (state == 1) {
            draw <- rbinom(1, 1, p[1, 2])
            if (draw == 1) {
                state <- 2 
            }
        } else {
            draw <- rbinom(1, 1, p[2, 1])
            if (draw == 1) {
                state <- 1
            }
        }
        states[i] <- state #I am assigning the ith value to be the state of the MCMC at the end of everyloop.
    }
    return(states)
}

# specify number of states
range <- c(1,2)

# plot the comparison
runs <- seq(1, 2000)
pct_results <- rep(NA, length(runs))
for (i in runs){
results <- mcmc(p= P_1, states, burnin=100, runs=i)
pct_results[i] <- sum(results[results==1]) / i 
}
Invariant = 1/3
plot(x=runs, y=pct_results, type = 'l', col = 'black', xlab = 'Iterations', ylab = 'State = 1', main = 'Convergence of the Markov Chain to the Invariant Distribution')
lines(Invariant, col = 'red', lwd= 5)

```

6.4.

```{r}

p <- 0.3
q <- 0.3
r <- 0.4

walk <- function(p, q, r, steps){
    outcome <- rep(NA, steps)
    place <- 0
    for (i in 1:steps){
        step <- which(rmultinom(1, 1, c(p, q, r))==1)
        if (step == 1) place <- place + 1
        if (step == 3) place <- place - 1
        outcome[i] <- place
    }
    return(outcome)
}

# run function
steps <- 1000
p <- 0.3
q <- 0.3
r <- 0.4
results <- walk(p, q, r, steps)

steps2 <- 1000
p2 <- 0.1
q2 <- 0.8
r2 <- 0.1
results2 <- walk(p2, q2, r2, steps2)

steps3 <- 1000
p3 <- 0.6
q3 <- 0.2
r3 <- 0.2
results3 <- walk(p3, q3, r3, steps3)

# plot function
plot(x=1:steps, y=results, type='l', xlab="step number", ylab="cumulative steps", main="Random Walk")
plot(x=1:steps2, y=results2, type='l', xlab="step number", ylab="cumulative steps", main="Random Walk")
plot(x=1:steps3, y=results3, type='l', xlab="step number", ylab="cumulative steps", main="Random Walk")

```


## PART IV

``` {r}

# Time until next patient arrives
patient_Arrival <- function() { round(rpois(n=1, lambda=10), 0) }
# Time spent with each patient
appointment_Length <- function() { round(runif(n=1, min=5, max=20), 0) }
# function checks whether a doctor is available
doc_Availability <- function(x) { return(which(x[1] == 0)) }
# construct time measure
calc_Minutes <- function(hours) { return(60 * hours) }
# intake walk-in patient
patient_Intake <- function(totalPatients, waitLengths, doctors, waitingRoom) {
    docsAvailable <- which( lapply(doctors, doc_Availability) == 1 )
    if (length(docsAvailable) > 0) {
        doc <- sample(docsAvailable, 1)
        doctors[[doc]]$patient <- 1
        doctors[[doc]]$apptlength <- appointment_Length()
        waitLengths <- c(waitLengths, 0)
    } else {
        waitingRoom <- rbind(c('patient', 0), waitingRoom)
    }
    totalPatients <- totalPatients + 1
    nextPatient <- patient_Arrival()
    return(list(totalPatients = totalPatients, nextPatient = nextPatient,
                waitLengths = waitLengths, doctors = doctors, waitingRoom = waitingRoom))
}
# time moves forward by 1 minute
time_Change <- function(time, waitLengths, doctors, minutes, waitingRoom) {
    # Countdown until next patient
    nextPatient <- nextPatient - time
    # Countdown for current appointments, reset
    # doctor if appointment ends
    for (i in 1:length(doctors)) {
        if (doctors[[i]]$apptlength != 0) {
            doctors[[i]]$apptlength <- doctors[[i]]$apptlength - time
        } else {
            doctors[[i]]$patient <- 0
        }
    }
   # Assign longest waiter to available doctor
  docsAvailable <- which( lapply(doctors, doc_Availability) == 1 )
    for (i in 1:nrow(waitingRoom)){
    if (length(docsAvailable) > 0 & nrow(waitingRoom) != 0) {
        doc <- sample(docsAvailable, 1)
        doctors[[doc]]$patient <- 1
        doctors[[doc]]$apptlength <- appointment_Length()
    # Choose the person with the longest wait time.
    waitNum <- which.max(waitingRoom[, 2])
    waitLengths <- c(waitLengths, max(as.numeric(waitingRoom[, 2])))
    waitingRoom <- waitingRoom[-waitNum,]
        }
    }
    # increase wait time by 1 minute for each person in wait room
    waitingRoom[, 2] <- as.numeric(waitingRoom[, 2]) + time
    # reduce remaining minutes for accepting new patients
    minutes <- minutes - time
    return(list(waitingRoom = waitingRoom, waitLengths = waitLengths,
                nextPatient = nextPatient, doctors = doctors,
                minutes = minutes))
}

# Main Loop
##############
# Create doctors
doctors <- list( list('patient' = 0, 'apptlength' = 0),
                 list('patient' = 0, 'apptlength' = 0),
                 list('patient' = 0, 'apptlength' = 0) )
# Create waiting room
waitingRoom <- data.frame(matrix(ncol=2, nrow=0))
colnames(waitingRoom) <- c('waiting_patients', 'wait_time')
# containers
totalPatients <- 0
waitLengths <- c()
minutes <- calc_Minutes(hours=7)
nextPatient <- 0
# main loop
while (length(which(lapply(doctors, doc_Availability) == 1)) != 3 |
        nrow(waitingRoom) != 0 | minutes > 0) {{
    if (nextPatient == 0 & minutes > 0) {
        intakeInfo <- patient_Intake(totalPatients, waitLengths, doctors, waitingRoom)
        nextPatient <- intakeInfo$nextPatient
        totalPatients <- intakeInfo$totalPatients
        waitLengths <- intakeInfo$waitLengths
        doctors <- intakeInfo$doctors
        waitingRoom <- intakeInfo$waitingRoom
        }
    timeInfo <- time_Change(time=1, waitLengths, doctors, minutes, waitingRoom)
    waitLengths <- timeInfo$waitLengths
    nextPatient <- timeInfo$nextPatient
    doctors <- timeInfo$doctors
    minutes <- timeInfo$minutes
    waitingRoom <- timeInfo$waitingRoom
}
results <- c(mean(waitLengths), mean(waitLengths[waitLengths != 0]), sum(waitLengths > 0), totalPatients, abs(minutes))
}
results

```

```{r}
# simulating main loop 1000 times:
trials <- 100
results <- data.frame(matrix(nrow=trials, ncol=5))
colnames(results) <- c('avg_wait', 'avg_wait2', 'count_wait', 'patient_count', 'minutes_past')
for (j in 1:trials) {
    doctors <- list( list('patient' = 0, 'apptlength' = 0),
                     list('patient' = 0, 'apptlength' = 0),
                     list('patient' = 0, 'apptlength' = 0) )
    waitingRoom <- data.frame(matrix(ncol=2, nrow=0))
    colnames(waitingRoom) <- c('waiting_patients', 'wait_time')
    totalPatients <- 0
    waitLengths <- c()
    minutes <- calc_Minutes(hours=7)
    nextPatient <- 0
  while (length(which(lapply(doctors, doc_Availability) == 1)) != 3 |
        nrow(waitingRoom) != 0 | minutes > 0) {
    if (nextPatient == 0 & minutes > 0) {
        intakeInfo <- patient_Intake(totalPatients, waitLengths, doctors, waitingRoom)
        nextPatient <- intakeInfo$nextPatient
        totalPatients <- intakeInfo$totalPatients
        waitLengths <- intakeInfo$waitLengths
        doctors <- intakeInfo$doctors
        waitingRoom <- intakeInfo$waitingRoom
        }
    timeInfo <- time_Change(time=1, waitLengths, doctors, minutes, waitingRoom)
    waitLengths <- timeInfo$waitLengths
    nextPatient <- timeInfo$nextPatient
    doctors <- timeInfo$doctors
    minutes <- timeInfo$minutes
    waitingRoom <- timeInfo$waitingRoom
  }
results[j,] <- c(mean(waitLengths), mean(waitLengths[waitLengths != 0]), sum(waitLengths > 0), totalPatients, abs(minutes))
}
results
quantile(waitLengths)
quantile(totalPatients)
quantile(minutes)


```

- If I change the paramters of lambda, I get different waiting times:

```{r}
# Time until next patient arrives
patient_Arrival <- function() { round(rpois(n=1, lambda=5), 0) }
# Time spent with each patient
appointment_Length <- function() { round(runif(n=1, min=5, max=20), 0) }
# function checks whether a doctor is available
doc_Availability <- function(x) { return(which(x[1] == 0)) }
# construct time measure
calc_Minutes <- function(hours) { return(60 * hours) }
# intake walk-in patient
patient_Intake <- function(totalPatients, waitLengths, doctors, waitingRoom) {
    docsAvailable <- which( lapply(doctors, doc_Availability) == 1 )
    if (length(docsAvailable) > 0) {
        doc <- sample(docsAvailable, 1)
        doctors[[doc]]$patient <- 1
        doctors[[doc]]$apptlength <- appointment_Length()
        waitLengths <- c(waitLengths, 0)
    } else {
        waitingRoom <- rbind(c('patient', 0), waitingRoom)
    }
    totalPatients <- totalPatients + 1
    nextPatient <- patient_Arrival()
    return(list(totalPatients = totalPatients, nextPatient = nextPatient,
                waitLengths = waitLengths, doctors = doctors, waitingRoom = waitingRoom))
}
# time moves forward by 1 minute
time_Change <- function(time, waitLengths, doctors, minutes, waitingRoom) {
    # Countdown until next patient
    nextPatient <- nextPatient - time
    # Countdown for current appointments, reset
    # doctor if appointment ends
    for (i in 1:length(doctors)) {
        if (doctors[[i]]$apptlength != 0) {
            doctors[[i]]$apptlength <- doctors[[i]]$apptlength - time
        } else {
            doctors[[i]]$patient <- 0
        }
    }
   # Assign longest waiter to available doctor
  docsAvailable <- which( lapply(doctors, doc_Availability) == 1 )
    for (i in 1:nrow(waitingRoom)){
    if (length(docsAvailable) > 0 & nrow(waitingRoom) != 0) {
        doc <- sample(docsAvailable, 1)
        doctors[[doc]]$patient <- 1
        doctors[[doc]]$apptlength <- appointment_Length()
    # Choose the person with the longest wait time.
    waitNum <- which.max(waitingRoom[, 2])
    waitLengths <- c(waitLengths, max(as.numeric(waitingRoom[, 2])))
    waitingRoom <- waitingRoom[-waitNum,]
        }
    }
    # increase wait time by 1 minute for each person in wait room
    waitingRoom[, 2] <- as.numeric(waitingRoom[, 2]) + time
    # reduce remaining minutes for accepting new patients
    minutes <- minutes - time
    return(list(waitingRoom = waitingRoom, waitLengths = waitLengths,
                nextPatient = nextPatient, doctors = doctors,
                minutes = minutes))
}
# Main Loop
##############
# Create doctors
doctors <- list( list('patient' = 0, 'apptlength' = 0),
                 list('patient' = 0, 'apptlength' = 0),
                 list('patient' = 0, 'apptlength' = 0) )
# Create waiting room
waitingRoom <- data.frame(matrix(ncol=2, nrow=0))
colnames(waitingRoom) <- c('waiting_patients', 'wait_time')
# containers
totalPatients <- 0
waitLengths <- c()
minutes <- calc_Minutes(hours=7)
nextPatient <- 0
# main loop
while (length(which(lapply(doctors, doc_Availability) == 1)) != 3 |
        nrow(waitingRoom) != 0 | minutes > 0) {{
    if (nextPatient == 0 & minutes > 0) {
        intakeInfo <- patient_Intake(totalPatients, waitLengths, doctors, waitingRoom)
        nextPatient <- intakeInfo$nextPatient
        totalPatients <- intakeInfo$totalPatients
        waitLengths <- intakeInfo$waitLengths
        doctors <- intakeInfo$doctors
        waitingRoom <- intakeInfo$waitingRoom
        }
    timeInfo <- time_Change(time=1, waitLengths, doctors, minutes, waitingRoom)
    waitLengths <- timeInfo$waitLengths
    nextPatient <- timeInfo$nextPatient
    doctors <- timeInfo$doctors
    minutes <- timeInfo$minutes
    waitingRoom <- timeInfo$waitingRoom
}
results <- c(mean(waitLengths), mean(waitLengths[waitLengths != 0]), sum(waitLengths > 0), totalPatients, abs(minutes))
}
results


```

```{r}

# simulating main loop 100 times for lamba =5
trials <- 100
results <- data.frame(matrix(nrow=trials, ncol=5))
colnames(results) <- c('avg_wait', 'avg_wait2', 'count_wait', 'patient_count', 'minutes_past')
for (j in 1:trials) {
    doctors <- list( list('patient' = 0, 'apptlength' = 0),
                     list('patient' = 0, 'apptlength' = 0),
                     list('patient' = 0, 'apptlength' = 0) )
    waitingRoom <- data.frame(matrix(ncol=2, nrow=0))
    colnames(waitingRoom) <- c('waiting_patients', 'wait_time')
    totalPatients <- 0
    waitLengths <- c()
    minutes <- calc_Minutes(hours=7)
    nextPatient <- 0
  while (length(which(lapply(doctors, doc_Availability) == 1)) != 3 |
        nrow(waitingRoom) != 0 | minutes > 0) {
    if (nextPatient == 0 & minutes > 0) {
        intakeInfo <- patient_Intake(totalPatients, waitLengths, doctors, waitingRoom)
        nextPatient <- intakeInfo$nextPatient
        totalPatients <- intakeInfo$totalPatients
        waitLengths <- intakeInfo$waitLengths
        doctors <- intakeInfo$doctors
        waitingRoom <- intakeInfo$waitingRoom
        }
    timeInfo <- time_Change(time=1, waitLengths, doctors, minutes, waitingRoom)
    waitLengths <- timeInfo$waitLengths
    nextPatient <- timeInfo$nextPatient
    doctors <- timeInfo$doctors
    minutes <- timeInfo$minutes
    waitingRoom <- timeInfo$waitingRoom
  }
results[j,] <- c(mean(waitLengths), mean(waitLengths[waitLengths != 0]), sum(waitLengths > 0), totalPatients, abs(minutes))
}
results
quantile(waitLengths)
quantile(totalPatients)
quantile(minutes)
```
